### 2013-10-23 main.tex

\documentclass{llncs}

% ======= Packages ======

%    ### General ###
%\usepackage{latexsym}
\usepackage{enumerate}

\usepackage{amssymb,amsmath}   %
%\usepackage{amsthm} % theorem environment

%    ### Format ###
%\usepackage{a4wide}     %% wide page
%\usepackage{times}      %% small fonts
%\usepackage{fullpage}

%    ### Logic and Theoretical Computer Science ###
\usepackage{stmaryrd}
\usepackage[all, knot]{xy}%%xy-pics
\usepackage{verbatim}
\usepackage[english]{babel}
\usepackage{graphicx}

\usepackage[table]{xcolor}

%    ### Game Theory ###

%\usepackage{xspace}
%\usepackage{tikz}                %avoid multiple call !!
  %\usetikzlibrary{arrows,shapes,automata,backgrounds,petri,fit,decorations.pathmorphing}

%\usepackage[active,tightpage]{preview}
%\PreviewEnvironment{tikzpicture}
%\setlength\PreviewBorder{5pt}

%\usepackage{lgcommands}
\usepackage{tikz,subfigure}
   \usetikzlibrary{shapes.geometric}

%    ### Linguistics ###
%\usepackage{tipa}   % International Phonetic Alphabet

%\usepackage{qtree}              %basicly   qtree and tikz-qtree do the samething
%\usepackage{tikz}                  % avoid multiple call !!
%\usepackage{tikz-qtree}

%\usepackage{gb4e}         % NOTE!!!   This should be the LAST  usepackage   call!!!!!!!
%\usepackage{lingmacros}

%%% Micha's stuff
\usepackage{booktabs}
\usepackage{todonotes}

% ======== Definitions ========

\newcommand{\intp}[1]{\llbracket #1 \rrbracket}         %interpretation

% ======== Title ========

\title{Variations on a Bayesian Theme: Comparing Bayesian Models of Referential Reasoning}
\author{Michael Franke \and Ciyang Qing}
\institute{Institute for Logic, Language and Computation}

% ======== Content ========

\begin{document}

\maketitle

\begin{abstract}
Recent developments in Bayesian experimental pragmatics have received much attention, as it provides us with a quantitative model formalizing core concepts in traditional pragmatic theories and the predictions it makes fit empirical data nicely. Despite its theoretical advantage and empirical success, there are still several conceptual issues in the Bayesian Rational Speech Act (RSA) model that need further clarification. In this paper, we analyze a few design choices made in RSA, focusing on its \emph{belief}, \emph{goal} and \emph{action} components and introduce some intuitively equally plausible alternatives motivated from related game-theoretic approaches to pragmatics as different interpretations of these three components. Then we compare the original RSA model with these alternatives to new empirical data and the result suggests a more complex picture of pragmatic reasoning about referential expressions than the RSA depicts. While there is a correlation between the speaker's and the listener's pragmatic reasoning that may be seen as common rationality, either side has its own bias that the other does not take into account.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\input{01_intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Referential Games and the RSA model}

\input{02_RSA}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alternatives in Design Choices}

\input{03_alternatives}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Experiment}

\input{04_experiment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model Comparison}

\input{05_model_comparison}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

\input{06_conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{splncs}
\bibliography{variations_ref}

\end{document}