### 2013-10-28 main.tex
revised abstract

\documentclass{llncs}

% ======= Packages ======
\usepackage{etex}
%    ### General ###
%\usepackage{latexsym}
\usepackage{enumerate}

\usepackage{amssymb,amsmath}   %
%\usepackage{amsthm} % theorem environment

%    ### Format ###
%\usepackage{a4wide}     %% wide page
%\usepackage{times}      %% small fonts
%\usepackage{fullpage}

%    ### Logic and Theoretical Computer Science ###
\usepackage{stmaryrd}
\usepackage[all, knot]{xy}%%xy-pics
\usepackage{verbatim}
\usepackage[english]{babel}
\usepackage{graphicx}

\usepackage[table]{xcolor}

%    ### Game Theory ###

%\usepackage{xspace}
%\usepackage{tikz}                %avoid multiple call !!
  %\usetikzlibrary{arrows,shapes,automata,backgrounds,petri,fit,decorations.pathmorphing}

%\usepackage[active,tightpage]{preview}
%\PreviewEnvironment{tikzpicture}
%\setlength\PreviewBorder{5pt}

%\usepackage{lgcommands}
\usepackage{tikz,subfigure}
   \usetikzlibrary{shapes.geometric}
\usepackage{pgfplots}
%    ### Linguistics ###
%\usepackage{tipa}   % International Phonetic Alphabet

%\usepackage{qtree}              %basicly   qtree and tikz-qtree do the samething
%\usepackage{tikz}                  % avoid multiple call !!
%\usepackage{tikz-qtree}

%\usepackage{gb4e}         % NOTE!!!   This should be the LAST  usepackage   call!!!!!!!
%\usepackage{lingmacros}

%%% Micha's stuff
\usepackage{booktabs}
\usepackage{todonotes}

% ======== Definitions ========

\newcommand{\intp}[1]{\llbracket #1 \rrbracket}         %interpretation

% ======== Title ========

\title{Variations on a Bayesian Theme: Comparing Bayesian Models of Referential Reasoning}
\author{Michael Franke \and Ciyang Qing}
\institute{Institute for Logic, Language and Computation}

% ======== Content ========

\begin{document}

\maketitle

\begin{abstract}
Recent developments in Bayesian experimental pragmatics have received much attention. The Rational Speech Act (RSA) model formalizes core concepts in traditional pragmatic theories quantitatively and makes predictions that fit empirical data nicely. In this paper, we analyze the RSA model and its relation to the closely related game-theoretic approaches to pragmatics, by spelling out its \emph{belief}, \emph{goal} and \emph{action} components. We introduce some alternatives motivated from the game-theoretic models and compare them with the RSA model to new experimental data. The result suggests that the RSA model could be adapted and extended to improve its predictive power.

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\input{01_intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Referential Games and the RSA model}

\input{02_RSA}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alternatives in Design Choices}

\input{03_alternatives}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Experiment}

\input{04_experiment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model Comparison}

\input{05_model_comparison}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

\input{06_conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{splncs}
\bibliography{variations_ref}

\end{document}