### 2013-10-30 main.tex

\documentclass{llncs}

% ======= Packages ======
\usepackage{etex}
%    ### General ###
%\usepackage{latexsym}
\usepackage{enumerate}

\usepackage{amssymb,amsmath}   %
%\usepackage{amsthm} % theorem environment

%    ### Format ###
%\usepackage{a4wide}     %% wide page
%\usepackage{times}      %% small fonts
%\usepackage{fullpage}

%    ### Logic and Theoretical Computer Science ###
\usepackage{stmaryrd}
\usepackage[all, knot]{xy}%%xy-pics
\usepackage{verbatim}
\usepackage[english]{babel}
\usepackage{graphicx}

\usepackage[table]{xcolor}

%    ### Game Theory ###

%\usepackage{xspace}
%\usepackage{tikz}                %avoid multiple call !!
  %\usetikzlibrary{arrows,shapes,automata,backgrounds,petri,fit,decorations.pathmorphing}

%\usepackage[active,tightpage]{preview}
%\PreviewEnvironment{tikzpicture}
%\setlength\PreviewBorder{5pt}

%\usepackage{lgcommands}
\usepackage{tikz,subfigure}
   \usetikzlibrary{shapes.geometric}
\usepackage{pgfplots}
%    ### Linguistics ###
%\usepackage{tipa}   % International Phonetic Alphabet

%\usepackage{qtree}              %basicly   qtree and tikz-qtree do the samething
%\usepackage{tikz}                  % avoid multiple call !!
%\usepackage{tikz-qtree}

%\usepackage{gb4e}         % NOTE!!!   This should be the LAST  usepackage   call!!!!!!!
%\usepackage{lingmacros}

%%% Micha's stuff
\usepackage{booktabs}
\usepackage{todonotes}

% ======== Definitions ========

\newcommand{\intp}[1]{\llbracket #1 \rrbracket}         %interpretation

% ======== Title ========

\title{Variations on a Bayesian Theme: Comparing Bayesian Models of
  Referential Reasoning}
\titlerunning{Variation on a Bayesian Theme}
\author{Ciyang Qing \and Michael Franke\thanks{Thanks to \dots % Judith Degen,
  % Noah Goodman, Daniel Lassiter [?], audience of the workshop, Henk
  % & Hans-Christian for organizing, reviewers \dots 
}}
\institute{Institute for Logic, Language and Computation\\Universtity
  of Amsterdam, The Netherlands}

% ======== Content ========

\begin{document}

\maketitle

\begin{abstract}
  Recent developments in Bayesian experimental pragmatics have
  received much attention. The Rational Speech Act (RSA) model
  formalizes core concepts of traditional pragmatic theories
  quantitatively and makes predictions that fit empirical data
  nicely. In this paper, we analyze the RSA model and its relation to
  closely related game theoretic approaches, by spelling out its
  \emph{belief}, \emph{goal} and \emph{action} components. We
  introduce some alternatives motivated from the game theoretic
  tradition and compare models incorporating these alternatives
  systematically to the original RSA model, using Bayesian model
  comparison, in terms of their ability to predict relevant empirical
  data. The result suggests that the RSA model could be adapted and
  extended to improve its predictive power, in particular by taking
  speaker preferences into account. 
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\input{01_intro}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Referential Games and the RSA model}
\label{sec:refer-games-rsa}

\input{02_RSA}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alternatives in Design Choices}
\label{sec:alternatives}

\input{03_alternatives}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Experiment}
\label{sec:experiment}

\input{04_experiment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model Comparison}
\label{sec:model-comparison}

\input{05_model_comparison}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:conclusion}

\input{06_conclusion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliographystyle{splncs}
\bibliography{variations_ref}

\end{document}