### 2013-10-30 02_RSA.tex
Sections 2,3 proof

Referential games are confined interactive reasoning tasks used to
study pragmatic inference in a controlled environment \cite{Frank}. A
referential game consists of a set of objects called a
\textit{context}. A context contains different shapes of different
colors in various arrangements, such as shown in Figure~\ref{context}.
The speaker in the game is asked to refer to one of the objects (the
\emph{target}) by uttering an expression (typically a single word
denoting a feature, i.e., color or shape, of the target) to the
listener. The listener, who has does not know which object is the
target, needs to recover the target based on the speaker's choice of
expression.

\begin{figure}[h] 
  \centering \subfigure[Example context: green square and circle, and
  blue circle.]{\label{context}
  \begin{tikzpicture}
     \path (-1.8, 0) node [shape=rectangle, draw=green!65 ,fill= green!65 ,minimum size=28]{}
           ( 0, 0) node  [shape=circle, draw=green!65, fill= green!65,minimum size=30] {}
           ( 1.8, 0) node [shape=circle, draw=blue!65, fill=blue!65  ,minimum size=30] {}
           ( -1.8, -0.8) node {$t_{1}$}
           ( -0.0, -0.8) node {$t_{2}$}
           (  1.8, -0.8) node {$t_{3}$}
     ;

  \end{tikzpicture}
}\hspace{5mm} \subfigure[Vocabulary and truth table]{\label{VaTT}
  \begin{tikzpicture}
     \path (0,0) node[] 
     {\begin{tabular}{r|ccc}
                         & \quad $t_{1}$\ \  &\quad $t_{2}$\ \  &\quad $t_{3}$ \ \   \\
    \hline
   ``Green'':  $m_{1}$ \quad  &  $1$        &   $1$       & $0$      \\
   ``Circle'': $m_{2}$  \quad &  $0$        &   $1$       & $1$      \\ 
   ``Square'': $m_3$ \quad    &  $1$      &   $0$       & $0$      \\
   ``Blue'':   $m_4$  \quad  &  $0$        &   $0$       & $1$     
     \end{tabular}
     } ;
  \end{tikzpicture}
  }
  \caption{A simple referential game}
\end{figure}

Let us first try to play this game ourselves, to get a better sense of
what it is about and why it is relevant to human pragmatic
reasoning. For simplicity let us assume that the utterances that the
speaker is allowed to use are from a \emph{vocabulary} which is
commonly known between the speaker and the listener. The content of
the vocabulary and the (truth-conditional/literal) meaning of each
word are shown in Figure \ref{VaTT}. Suppose we play as the speaker
and the target is $t_1$, the green square. We can either use
``Square'' or ``Green'' to truthfully refer to it, but it seems more
prudent to use ``Square'', as there is only one square in the context
and thus the listener can easily identify it, whereas there are two
green objects, which makes ``Green'' ambiguous. In terms of the
Gricean \emph{Maxim of Quantity}, we should use ``Square'' because it
is more informative than ``Green'' in the given context; the surplus
informativity moreover seems relevant. Similarly, we should use
``Blue'' to refer to $t_3$, the blue circle. However, using ``Blue''
to refer to the blue circle might intuitively sound a little
unnatural, as color terms are usually used as adjectives\footnote{They
  are used as nouns mostly to refer to the colors themselves.}, while
we usually use nouns to refer to concrete objects. This inclination
becomes more evident when we want to refer to $t_2$, the green
circle. While ``Green'' and ``Circle'' are equally ambiguous in this
case, we might nevertheless prefer the latter.

Now let us turn to play as the listener. If we hear ``Square'' or
``Blue'', we easily know the intended referent as there is no
ambiguity, but what if we hear ``Circle'' (or ``Green'')? There are
two circles in the context that we need to choose from. On the one
hand, the blue circle, having the unique color in the context, seems
to be perceptually dominant and thus easily captures our attention. On
the other hand, from the previous analysis we know that if the blue
circle were the intended referent, the speaker could have chosen
``Blue'' which is not ambiguous and thus more informative. Hence the
listener needs to balance two sources of information, i.e., the
(presumably subconscious) perceptual salience of different objects and
the rational expectation of the likelihood of the speaker making the
utterance for each object. The latter line of reasoning is crucial to
the classic Gricean account of scalar implicatures, and the major
challenges in terms of formal modeling are how to quantify notions
such as informativeness and how different pieces of information should
be integrated.

The RSA model addresses the problems by using information-theoretic
concepts to measure informativeness and Bayesian inference to
integrate different sources of information
\cite{Frank2009,Frank,Bergen2012,GoodmanStuhlmuller2013:Knowledge-and-I}. In
order to measure the informativeness of an utterance, the RSA model
starts with the literal listener who upon receiving an utterance $m$
does a conditioning on its literal meaning:
\begin{equation}\label{Bayesian_literal}
\rho_0(t \mid m) = \mathcal{U}(t\mid \intp{m}) ,
\end{equation}
where $\mathcal{U}$ is a uniform distribution over all the possible
referents. The informativeness of utterance $m$ for the intended
referent $t$ can be measured as the negative Kullback-Leibler
divergence of the induced literal listener's belief $\rho_0$ from the
speaker's own belief $\delta_t$:
\begin{equation} \label{Bayesian_sender_informativeness}
\mbox{Info}(m,t)=-\mbox{KL}(\delta_t \| \rho_0)=-\sum_{t'}\delta_t(t')\log\left(\frac{\delta_t(t')}{\rho_0(t')}\right),
\end{equation}
where $\delta_t$ is a delta distribution with all probability mass on
target object $t$, as the speaker knows her intended referent:
\begin{equation} \label{Bayesian_sender_delta}
\delta_t(t')=\left\{ \begin{array}{ll}
1 & \mbox{if } t' = t \\
0 & \mbox{otherwise}
\end{array}\right. \enspace .
\end{equation}

The speaker acts sub-optimally by choosing the utterance that soft-maximizes her expected utility, which is defined as the informativeness of the utterance subtracted by its cost:
\begin{equation} \label{Bayesian_sender_softmax}
\sigma(m \mid t) \propto \exp(\lambda_\mathrm{S}\cdot \mbox{U}(m,t))=\exp(\lambda_\mathrm{S} \cdot (\mbox{Info}(m,t)-\mbox{Cost}(m))),
\end{equation}
where $\lambda_\mathrm{S}$ is a parameter measuring the speaker's
degree of rationality, i.e., to what extent the speaker sticks to the
strict optimum. The cost term is used to encode preference in
different utterances, be it about the utterances' lengths or syntactic
categories.

From (\ref{Bayesian_literal})-(\ref{Bayesian_sender_softmax}) we obtain the speaker's production rule:

\begin{equation} \label{Bayesian_speaker}
\sigma(m \mid t) \propto \exp(\lambda_\mathrm{S} \cdot (\log \mathcal{U}(t\mid \intp{m})-\mbox{Cost}(m))) \enspace .
\end{equation}

The pragmatic listener in the RSA model, upon receiving the utterance
$m$, performs a Bayesian update on his prior belief $\mathcal{S}(t)$
by using an estimate of the speaker's behavior
(\ref{Bayesian_speaker}):
\begin{equation} \label{Bayesian_rec_update}
\rho(t \mid m) \propto \mathcal{S}(t)\cdot \sigma(m \mid t) \enspace .  
\end{equation}
Bayes' rule naturally integrates the perceptual salience of each
object, which is treated as the prior $\mathcal{S}(t)$ and can be
empirically measured, with listener's expectation of the speaker being
informative, which is incorporated as the likelihood, thus addressing
the previously mentioned challenge of balancing different sources of
information. Setting $\lambda=1$ and $\mbox{Cost}(m)=0$ for all $m$,
\cite{Frank} obtained a highly significant correlation between the
model prediction and their experiment data on actual speaker and
listener behavior gathered from referential games of varying
complexity.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% TeX-PDF-mode: t
%%% End: