### 2013-10-30 06_conclusion.tex

Our experiment and model comparison suggest the following: (1) On the
conceptual level, it is helpful to clarify the distinction between
belief, goal and action, and the Bayesian framework provides us with a
natural way to structure these components into a formal model. By
examining each component systematically, we can explicitly spell out
the underlying assumptions and formulate alternative hypotheses to be
tested. (2) On the empirical level, we tested the RSA model with a
family of variants motivated from the game theoretic pragmatic
perspectives on communication, and in particular, we investigated the
speaker's preference as well as the listener's perceptual
salience. Our data showed a more intricate picture about the roles
various factors play in pragmatic reasoning about referential
expressions. There is some correlation between the speaker's and the
listener's rationality, while either side has his own bias and is
unaware of the other's. Also, an action-based view might better
reflect the goal of communication, at least in the forced-choice
experimental setting. Understanding the relation between the
forced-choice and the betting paradigms would be an important next
step.

Our work shows that careful conceptual analysis of the design choices for quantitative models can lead to a better understanding of the phenomenon and further improvement in the formal model's predictive power. Of course, our preliminary experiment should only serve as a start and more empirical data are needed for a robust pragmatic theory of people's reasoning about referential expressions.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% TeX-PDF-mode: t
%%% End: