### 2013-10-30 06_conclusion.tex

Our experiment and model comparison suggest the following: (1) On the
conceptual level, it is helpful to clarify the distinction between
belief, goal and action, and the Bayesian framework provides us with a
natural way to structure these components into a formal model. By
examining each component systematically, we can explicitly spell out
the underlying assumptions and formulate alternative hypotheses to be
tested. (2) On the empirical level, we tested the RSA model with a
family of variants motivated from a game theoretic perspective on
communication, and in particular, we investigated the speaker's
preference as well as the listener's perceptual salience. Our data
showed a more intricate picture of the roles that various factors play
in pragmatic reasoning about referential expressions. We found
evidence for a correlation between the speaker's and the listener's
rationality. Either side appears to have his own biases, but appears
to be negligent of the other's. Also, an action-based view might
better reflect the goal of communication, at least in the
forced-choice experimental setting. Understanding the relation between
the forced-choice and the betting paradigms would be an important next
step to gain more insight into the differences between action- and
belief-based goal structures.

Our work shows that careful conceptual analysis of the design choices
for quantitative models can lead to a better understanding of the
phenomenon and further improvement in the formal model's predictive
power. Of course, our restricted empirical data can only serve as a
start and more data is needed to fuel further model comparison towards
a more robust pragmatic theory of people's probabilistic reasoning
about referential expressions.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% TeX-PDF-mode: t
%%% End: